The perceptual dimensions of \\ sonority-driven epenthesis

### 2013-10-09 paper.tex
Lots of minor changes

Vowel epenthesis often appears to preferentially target consonant clusters with rising sonority.
There are two broad classes of explanation within Optimality Theory for such sonority-driven epenthesis.

One is faithfulness-based: the perceptual distance between the underlying /C\textsubscript{1}C\textsubscript{2}/ sequence and its correspondent
output sequence [C\textsubscript{1}VC\textsubscript{2}] is small when the cluster is of rising sonority.
Thus, epenthesis into such a sequence incurs a smaller faithfulness cost than epenthesis into a cluster of falling sonority. This is the basis of the analysis proposed by Fleischhacker (2002, 2005) to explain why rising sonority obstruent-sonorant clusters are more easily epenthesised in
to than falling sonority sibilant-stop clusters.

This faithfulness-based approach raises the question of how the perceptual distance between two sonority contours /C\textsubscript{1}C\textsubscript{2}/ and [C\textsubscript{1}VC\textsubscript{2}] should be computed in terms of the sonority of C\textsubscript{1}, C\textsubscript{2} and V.

@@

The layout of this paper is as follows.  Section \ref{theoreticalmachinery} lays out the theoretical background for the sonority contour faithfulness approach to 
sonority-driven epenthesis. I introduce the proposed {\sc Sonority Angle} metric as well as the competing {\sc Sonority Rise} metric \citep{flemming.2008}, 
then lay out the alternative markedness-based approach to sonority-driven epenthesis, namely Syllable Contact.

@@

Section \ref{chaha} is a case study of epenthesis positioning in Chaha.  I detail the facts of epenthesis positioning in Chaha, based on the data given in \citet{rose.2000},
and show that the sonority contour faithfulness approach explains these facts, with {\sc Sonority Angle} as the metric for comparing sonority contours.  I compare it to {\sc Sonority Rise} and show that the former is the more successful analysis, and that overall, the approach just outlined is more economical than the  Syllable Contact-based approach of \citet{rose.2000}.

Section \ref{issues} looks at various issues regarding {\sc Sonority Angle}, such as its robustness. Section \ref{conclusion} concludes.

@@

The second generalisation is that the more sonorous the C\textsubscript{1}, the more likely
epenthesis is to occur. Let us fix the sonority distance between the two consonants of the cluster at 2, but vary their distance from V:

@@

There are two possible ways to rank the family of {\sc Ident(Son$\measuredangle$)}$<$n constraints.
One is to fix them in a universal ranking \citep{prince.smolensky.1993}, with least stringent the highest-ranked:

@@

\ex. Possible free ranking:
     {\sc Ident(Son$\measuredangle$)}$<$1.5 $\gg$ {\sc Ident(Son$\measuredangle$)}$<$2.0 $\gg$ ...

Depending on what constraints intervene, this may result in conflation of certain areas of the hierarchy, so that any distinctions between {\sc Sonority Angles} between 1.5 and 2.0 are made irrelevant.

@@

\ex. Ident(Son$\measuredangle$)$<$1.5 $\gg$ {\sc *ComplexCoda} $\gg$ {\sc Dep}
     \a. Allows epenthesis in rising and level clusters such as stop-fricative TF:
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/TF/} & {\sc Ident(Son$\measuredangle$)}$<$1.5 & {\sc *ComplexCoda} & {\sc Dep} \\[0.5ex]
\hline \hline a. & & TF & & $\ast$! & \cellcolor{lightgray} \\
\hline b. & \ding{43} & T\textipa{@}F & 0.59 & & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \b. Prevents epenthesis in falling clusters such as FT:
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/FT/} & {\sc Ident(Son$\measuredangle$)}$<$1.5 & {\sc *ComplexCoda} & {\sc Dep} \\[0.5ex]
\hline \hline c. & \ding{43} & FT & &\cellcolor{lightgray} $\ast$ & \cellcolor{lightgray} \\
\hline d. & & F\textipa{@}T & 2.11 $\ast$! &\cellcolor{lightgray} & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\ex. {\sc *ComplexCoda} $\gg$ {\sc Max}(r) $\gg$ {\sc Dep}

\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c:c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{hOrn}/} & {\sc *ComplexCoda} & {\sc Max}(r) & {\sc Dep} \\[0.5ex]
\hline \hline a. & & \textipa{hOrn} & $\ast$! & & \cellcolor{lightgray} \\
\hline b. & \ding{43} & \textipa{hOrOn} & & & \cellcolor{lightgray}$\ast$ \\
\hline c. & & \textipa{hOn} & & $\ast$! & \cellcolor{lightgray} \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\ex. {\sc *ComplexCoda}, {\sc Ident(Son$\measuredangle$)}$<$2.3  $\gg$ {\sc Max}(r) $\gg$ {\sc Dep}

\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c:c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{park}/} & {\sc *ComplexCoda} & {\sc Ident(Son$\measuredangle$)}$<$2.3 & {\sc Max}(r) & {\sc Dep} \\[0.5ex]
\hline \hline a. & & \textipa{park} & $\ast$! & & \cellcolor{lightgray} & \cellcolor{lightgray} \\
\hline b. & & \textipa{parak} & & $\ast$! & \cellcolor{lightgray} & \cellcolor{lightgray}$\ast$ \\
\hline c. & \ding{43} & \textipa{pak} & & & \cellcolor{lightgray}$\ast$ & \cellcolor{lightgray} \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

While other analyses could possibly account for this data (see one such in \citep{karim.2011}), we see that simply ranking {\sc Ident(Son$\measuredangle$)}$<$n $\gg$ $\mathcal{F}$ allows us to account for it quite straightforwardly.

@@

\bigskip
%\begin{center} \renewcommand*\arraystretch{1.2}
%\scalebox{1}[1]
{\begin{tabular}[t]{|rrl||c|c|c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/FT/} & {\sc Id(SA)}$<$2.5 & {\sc Id(SA)}$<$1.5 & {\sc *CompCoda} & {\sc Id(SA)}$<$1.0 & {\sc Id(SA)}$<$0.5 \\[0.5ex]
\hline \hline c. & \ding{43}& FT & & & \cellcolor{lightgray} $\ast$ & \cellcolor{lightgray} & \cellcolor{lightgray} \\
\hline d. &  & F\textipa{@}T & 2.11 & 2.11 $\ast$! & \cellcolor{lightgray} & \cellcolor{lightgray} 2.11 $\ast$ & \cellcolor{lightgray}2.11 $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\ex. {\sc Ident(Son$\measuredangle$)}$<$n $\gg$ $\mathcal{M}$
     \a. Scenario 1: {\sc SonAngle}(C\textsubscript{1}C\textsubscript{2}), {\sc SonAngle}(C\textsubscript{2},C\textsubscript{3})$<$n \\
         Since neither violates {\sc Ident(Son$\measuredangle$)}$<$n, $\mathcal{M}$ determines output.
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n & $\mathcal{M}$ \\[0.5ex]
\hline \hline a. & \ding{43} & C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $<$n & \\
\hline b. & & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $<$n & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
      \b. Scenario 2: {\sc SonAngle}(C\textsubscript{1}C\textsubscript{2})$>$n, {\sc SonAngle}(C\textsubscript{2},C\textsubscript{3})$<$n \\
          C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} harmonically bound by C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} candidate.
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline c. & \ding{43} &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3}  & $<$n & \cellcolor{lightgray} \\
\hline d. & & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $>$n $\ast$! & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \c. Scenario 3: {\sc SonAngle}(C\textsubscript{1}C\textsubscript{2})$<$n, {\sc SonAngle}(C\textsubscript{2},C\textsubscript{3})$>$n \\
         {\sc Ident(Son$\measuredangle$)}$<$n determines output, overruling $\mathcal{M}$.
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline e. & &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $>$n $\ast$!  & \cellcolor{lightgray} \\
\hline f. & \ding{43} & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3}& $<$n  & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \d. Scenario 4: {\sc SonAngle}(C\textsubscript{1}C\textsubscript{2}), {\sc SonAngle}(C\textsubscript{2}C\textsubscript{3}) $>$ n \\
         Since both candidates violate {\sc Ident(Son$\measuredangle$)}$<$n, $\mathcal{M}$ determines output.
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline g. & &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $>$n $\ast$  & \\
\hline h. & \ding{43} & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $>$n $\ast$ & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

@@

\begin{center}
\begin{tabular}{ccccccccccccc}
 V & $>$ & w j & $>$ & r & $>$ & \textipa{B} & $>$ & m n & $>$ & f s z x & $>$ & t t' k k' d g \\
 vowels & & glides & & liquids & &   & & nasals & & fricatives & & stops \\
    V   & &    G   & &    R    & & \textipa{B} & & N & & F & & T \\
    6   & &    5   & &    4    & &    3.5      & & 3 & & 2 & & 1 \\
\end{tabular}
\end{center}

@@

	\node (TR) at (0.124354994547*8, 1) {TR};
	\node (Tb) at (0.183110817262*8, 0.5) {T\textipa{B}};
	\node (FR) at (0.218668945874*8, 1) {FR};
	\node (TN) at (0.266252049151*8, 1.5) {TN};
	\node (Fb) at (0.343023940421*8, 1) {F\textipa{B}};
	\node (NR) at (0.463647609001*8, 1) {NR};
	\node (FN) at (0.540419500271*8, 0.5) {FN};
	\node (TF) at (0.588002603548*8, 1) {TF};
	\node (bR) at (0.726642340682*8, 0.5) {\textipa{B}R};
	\node (Nb) at (0.785398163397*8, 1) {N\textipa{B}};
	\node (RR) at (1.10714871779*8, 1) {RR};
	\node (bb) at (1.19028994968*8, 0.5) {\textipa{BB}};
	\node (NN) at (1.2490457724*8, 1) {NN};
	\node (FF) at (1.32581766367*8, 1.5) {FF};
	\node (TT) at (1.37340076695*8, 1) {TT};
	\node (Rb) at (1.57079632679*8, 1) {R\textipa{B}};
	\node (bN) at (1.65393755868*8, 0.5) {\textipa{B}N};
	\node (RN) at (1.89254688119*8, 1) {RN};
	\node (NF) at (2.0344439358*8, 1) {NF};
%	\node (GF) at (2.0344439358*8, 1.5) {GF};
	\node (FT) at (2.11121582707*8, 0.5) {FT};
	\node (bF) at (2.17308367293*8, 1) {\textipa{B}F};
	\node (RF) at (2.21429743559*8, 1.5) {RF};
	\node (RT) at (2.35619449019*8, 1) {RT};
	\node (NT) at (2.35619449019*8, 0.5) {NT};
	\node (bT) at (2.38057989937*8, 1.5) {\textipa{B}T};
  \node (rising) at (0.3*8,2.3) {\underline{Rising sonority}};
  \node (level) at (1.2*8,2.3) {\underline{Level sonority}};
  \node (falling) at (2.2*8,2.3) {\underline{Falling sonority}};
  % axis
  \node (axisstart) at (-0.23,0) {};
  \node (axisend)   at (2.7*8,0) {};
  \draw (axisstart) -- (axisend);
  \draw (0.0*8, 0.2) -- (0.0*8, -0.2) -- cycle;
  \draw (1.0*8, 0.2) -- (1.0*8, -0.2) -- cycle;
  \draw (2.0*8, 0.2) -- (2.0*8, -0.2) -- cycle;
  \node (0pointlabel) at (0.0,-0.5) {0};
  \node (1pointlabel) at (1.0*8,-0.5) {1};
  \node (2pointlabel) at (2.0*8,-0.5) {2};
  \node (xaxislabel) at (2.7*8,-0.5) {\textsc{SonAngle}};
  % dividing line
  \node (dividinglinestart) at (1.5 * 8,2) {};
  \node (dividinglineend)   at (1.5 * 8,-0.3) {};
  \path [line,dashed] (dividinglinestart) -- (dividinglineend) -- cycle;
  \node (dividingpointlabel) at (1.5*8, -0.5) {1.5};
\end{tikzpicture}
}

@@

The only clusters that give us evidence of this are /\textipa{Bm}/ clusters. These are not seen root-internally due to a high-ranking OCP effect \citep[cited in \citep{rose.2000}]{greenberg.1950}.
I suggest that there might be an undominated markedness constraint against this cluster -- either OCP(labial) or another constraint based on articulatory or perceptual difficulty -- that outranks all relevant faithfulness constraints.

@@

\ex. {\sc *CCC} $\gg$ {\sc Dep}

\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/CCC/} & *CCC & {\sc Dep} \\[0.5ex]
\hline \hline a. & & CCC & $\ast$! &  \\
\hline b. & \ding{43} & CC\textipa{1}C & & $\ast$ \\
\hline c. & \ding{43} & C\textipa{1}CC & & $\ast$ \\
\hline d. & & C\textipa{1}C\textipa{1}C & & $\ast\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\noindent {\bf [C\textipa{1}CC] idiolect.} In the third idiolect \ref{C1CCidiolect} (c.f. Section 5.2 of of \citep{rose.2000}), most clusters epenthesise as [C\textipa{1}CC].

@@

Several possibilities are possible: retain the sonority of the underlying C\textsubscript{1} for that of the output value; use the sonority of the output C\textsubscript{1} as the underlying value; or take the angle formed by the meeting of the two contours wherever it occurs.

>>>>

Vowel epenthesis often appears to preferentially target consonant clusters with rising sonority, as in the following example of English loanwords borrowed into Hawai'ian creole. The rising sonority onset cluster is adapted with epenthesis, while the falling sonority cluster is left intact.

\ex. \a. puranti: `plenty'
    \b. ste: `stay'
    \z.
    (\cite{nagara.1972}, cited by \cite{fleischhacker.2005})

There are two broad classes of explanation within Optimality Theory for such sonority-driven epenthesis.

One is faithfulness-based: the perceptual distance between the underlying /C\textsubscript{1}C\textsubscript{2}/ sequence and its correspondent
output sequence [C\textsubscript{1}VC\textsubscript{2}] is small when the cluster is of rising sonority.
Thus, epenthesis into such a sequence incurs a smaller faithfulness cost than epenthesis into a cluster of falling sonority. This is the basis of the analysis proposed by \cite{fleischhacker.2002, fleischhacker.2005} to explain why rising sonority obstruent-sonorant clusters are more easily epenthesised in
to than falling sonority sibilant-stop clusters.

This raises the question of how the perceptual distance between two sonority contours /C\textsubscript{1}C\textsubscript{2}/ and [C\textsubscript{1}VC\textsubscript{2}] should be computed in terms of the sonority of C\textsubscript{1}, C\textsubscript{2} and V.

@@

The layout of this paper is as follows.  Section \ref{theoreticalmachinery} lays out the theoretical background for the sonority contour faithfulness approach to 
sonority-driven epenthesis. I introduce the proposed {\sc Sonority Angle} metric as well as the competing {\sc Sonority Rise} metric \citep{flemming.2008}, 
then lay out the alternative markedness-based approach to sonority-driven epenthesis.

@@

Section \ref{chaha} is a case study of epenthesis positioning in Chaha.  I detail the facts of epenthesis positioning in Chaha, based on data from \cite{rose.2000},
and show that the sonority contour faithfulness approach explains these facts, with {\sc Sonority Angle} as the metric for comparing sonority contours.  I compare it to {\sc Sonority Rise} and show that the former is the more successful analysis, and that overall, the approach just outlined is comparably successful to the Syllable Contact-based approach of \citet{rose.2000}.

Section \ref{issues} discusses various issues regarding {\sc Sonority Angle}, such as its robustness. Section \ref{conclusion} concludes.

@@

The second generalisation is that the more sonorous the C\textsubscript{1}, the more likely
epenthesis is to occur. Let us fix the sonority distance between the two consonants of the cluster at 2, but vary their distance from V. Decreasing this distance decreases the {\sc Sonority Angle}.

@@

There are two possible ways to rank the family of {\sc Ident(Son$\measuredangle$)}$<$n constraints.
One is to fix them in a universal ranking \citep{prince.smolensky.1993}, with the least stringent constraint the highest-ranked:

@@

\ex. Possible free ranking: \label{freerankingex}
     {\sc Ident(Son$\measuredangle$)}$<$1.5 $\gg$ {\sc Ident(Son$\measuredangle$)}$<$2.0 $\gg$ ...

This represents a slight deviation from the P-map hypothesis \citep{steriade.2001} in implementation, though not in spirit. Because the constraints form a stringency hierarchy, there cannot be reversals of the cluster hierarchy. What may occur is the {\it conflation} of certain areas of the constraint hierarchy, such as all the clusters with {\sc Sonority Angles} between 1.5 and 2.0 in \ref{freerankingex}, depending on what constraints intervene. Some illustrative examples are given below.

@@

\ex. Ident(Son$\measuredangle$)$<$1.5 $\gg$ {\sc *ComplexCoda} $\gg$ {\sc Dep}
     \a. Allows epenthesis in rising and level clusters such as stop-fricative TF:
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/TF/} & {\sc Ident(Son$\measuredangle$)}$<$1.5 & {\sc *ComplexCoda} & {\sc Dep} \\[0.5ex]
\hline \hline a. & & TF & & $\ast$! & \cellcolor{lightgray} \\
\hline b. & \ding{43} & T\textipa{@}F & 0.59 & & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
\vspace{0.5em}
     \b. Prevents epenthesis in falling clusters such as FT:
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/FT/} & {\sc Ident(Son$\measuredangle$)}$<$1.5 & {\sc *ComplexCoda} & {\sc Dep} \\[0.5ex]
\hline \hline c. & \ding{43} & FT & &\cellcolor{lightgray} $\ast$ & \cellcolor{lightgray} \\
\hline d. & & F\textipa{@}T & 2.11 $\ast$! &\cellcolor{lightgray} & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\ex. {\sc *ComplexCoda} $\gg$ {\sc Max}(r) $\gg$ {\sc Dep}
\vspace{-1em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c:c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{hOrn}/} & {\sc *ComplexCoda} & {\sc Max}(r) & {\sc Dep} \\[0.5ex]
\hline \hline a. & & \textipa{hOrn} & $\ast$! & & \cellcolor{lightgray} \\
\hline b. & \ding{43} & \textipa{hOrOn} & & & \cellcolor{lightgray}$\ast$ \\
\hline c. & & \textipa{hOn} & & $\ast$! & \cellcolor{lightgray} \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\ex. {\sc *ComplexCoda}, {\sc Ident(Son$\measuredangle$)}$<$2.3  $\gg$ {\sc Max}(r) $\gg$ {\sc Dep}
\vspace{-1em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c:c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{park}/} & {\sc *ComplexCoda} & {\sc Ident(Son$\measuredangle$)}$<$2.3 & {\sc Max}(r) & {\sc Dep} \\[0.5ex]
\hline \hline a. & & \textipa{park} & $\ast$! & & \cellcolor{lightgray} & \cellcolor{lightgray} \\
\hline b. & & \textipa{parak} & & $\ast$! & \cellcolor{lightgray} & \cellcolor{lightgray}$\ast$ \\
\hline c. & \ding{43} & \textipa{pak} & & & \cellcolor{lightgray}$\ast$ & \cellcolor{lightgray} \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

While other analyses could possibly account for this data (see one such in \cite{karim.2011}), we see that simply ranking {\sc Ident(Son$\measuredangle$)}$<$n $\gg$ $\mathcal{F}$ allows us to account for it quite straightforwardly.

@@

\bigskip
%\begin{center} \renewcommand*\arraystretch{1.2}
%\scalebox{1}[1]
{\begin{tabular}[t]{|rrl||c|c|c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/FT/} & {\sc Id(SA)}$<$2.5 & {\sc Id(SA)}$<$1.5 & {\sc *CompCoda} & {\sc Id(SA)}$<$1.0 & {\sc Id(SA)}$<$0.5 \\[0.5ex]
\hline \hline c. & \ding{43}& FT & & & \cellcolor{lightgray} $\ast$ & \cellcolor{lightgray} & \cellcolor{lightgray} \\
\hline d. &  & F\textipa{@}T & 2.11 & 2.11 $\ast$! & \cellcolor{lightgray} & \cellcolor{lightgray} 2.11 $\ast$ & \cellcolor{lightgray}2.11 $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
\vspace{0.5em}

@@

\ex. {\sc Ident(Son$\measuredangle$)}$<$n $\gg$ $\mathcal{M}$
     \a. Scenario 1: {\sc SonAngle}(C\textsubscript{1}C\textsubscript{2}), {\sc SonAngle}(C\textsubscript{2},C\textsubscript{3})$<$n \\
         Since neither violates {\sc Ident(Son$\measuredangle$)}$<$n, $\mathcal{M}$ determines output.
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n & $\mathcal{M}$ \\[0.5ex]
\hline \hline a. & \ding{43} & C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $<$n & \\
\hline b. & & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $<$n & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
\vspace{0.5em}
      \b. Scenario 2: {\sc SonAngle}(C\textsubscript{1}C\textsubscript{2})$>$n, {\sc SonAngle}(C\textsubscript{2},C\textsubscript{3})$<$n \\
          C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} harmonically bound by C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} candidate.
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline c. & \ding{43} &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3}  & $<$n & \cellcolor{lightgray} \\
\hline d. & & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $>$n $\ast$! & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
\vspace{0.5em} \newpage
     \c. Scenario 3: {\sc SonAngle}(C\textsubscript{1}C\textsubscript{2})$<$n, {\sc SonAngle}(C\textsubscript{2},C\textsubscript{3})$>$n \\
         {\sc Ident(Son$\measuredangle$)}$<$n determines output, overruling $\mathcal{M}$.
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline e. & &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $>$n $\ast$!  & \cellcolor{lightgray} \\
\hline f. & \ding{43} & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3}& $<$n  & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
\vspace{0.5em}
     \d. Scenario 4: {\sc SonAngle}(C\textsubscript{1}C\textsubscript{2}), {\sc SonAngle}(C\textsubscript{2}C\textsubscript{3}) $>$ n \\
         Since both candidates violate {\sc Ident(Son$\measuredangle$)}$<$n, $\mathcal{M}$ determines output.
\vspace{-0.5em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C\textsubscript{1}C\textsubscript{2}C\textsubscript{3}/} & {\sc Ident(Son$\measuredangle$)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline g. & &  C\textsubscript{1}C\textsubscript{2}\textipa{@}C\textsubscript{3} & $>$n $\ast$  & \\
\hline h. & \ding{43} & C\textsubscript{1}\textipa{@}C\textsubscript{2}C\textsubscript{3} & $>$n $\ast$ & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
\vspace{0.5em}

@@

\vspace{-1em}

@@

\vspace{-1.5em}
\begin{center}
\begin{tabular}{ccccccccccccc}
 V & $>$ & w j & $>$ & r & $>$ & \textipa{B} & $>$ & m n & $>$ & f s z x & $>$ & t t' k k' d g \\
 vowels & & glides & & liquids & &   & & nasals & & fricatives & & stops \\
    V   & &    G   & &    R    & & \textipa{B} & & N & & F & & T \\
    6   & &    5   & &    4    & &    3.5      & & 3 & & 2 & & 1 \\
\end{tabular}
\end{center}

@@

	\node (TR) at (0.124354994547*8, 1) {TR};
	\node (Tb) at (0.183110817262*8, 0.5) {T\textipa{B}};
	\node (FR) at (0.218668945874*8, 1) {FR};
	\node (TN) at (0.266252049151*8, 1.5) {TN};
	\node (Fb) at (0.343023940421*8, 1) {F\textipa{B}};
	\node (NR) at (0.463647609001*8, 1) {NR};
	\node (FN) at (0.540419500271*8, 0.5) {FN};
	\node (TF) at (0.588002603548*8, 1) {TF};
	\node (bR) at (0.726642340682*8, 0.5) {\textipa{B}R};
	\node (Nb) at (0.785398163397*8, 1) {N\textipa{B}};
	\node (RR) at (1.10714871779*8, 1) {RR};
%	\node (bb) at (1.19028994968*8, 0.5) {\textipa{BB}};
	\node (NN) at (1.2490457724*8, 1) {NN};
	\node (FF) at (1.32581766367*8, 1.5) {FF};
	\node (TT) at (1.37340076695*8, 1) {TT};
	\node (Rb) at (1.57079632679*8, 1) {R\textipa{B}};
	\node (bN) at (1.65393755868*8, 0.5) {\textipa{B}N};
	\node (RN) at (1.89254688119*8, 1) {RN};
	\node (NF) at (2.0344439358*8, 1) {NF};
%	\node (GF) at (2.0344439358*8, 1.5) {GF};
	\node (FT) at (2.11121582707*8, 0.5) {FT};
	\node (bF) at (2.17308367293*8, 1) {\textipa{B}F};
	\node (RF) at (2.21429743559*8, 1.5) {RF};
	\node (RT) at (2.35619449019*8, 1) {RT};
	\node (NT) at (2.35619449019*8, 0.5) {NT};
	\node (bT) at (2.38057989937*8, 1.5) {\textipa{B}T};
  \node (rising) at (0.3*8,2.3) {\underline{Rising sonority}};
  \node (level) at (1.2*8,2.3) {\underline{Level sonority}};
  \node (falling) at (2.2*8,2.3) {\underline{Falling sonority}};
  % axis
  \node (axisstart) at (-0.23,0) {};
  \node (axisend)   at (2.7*8,0) {};
  \draw (axisstart) -- (axisend);
  \draw (0.0*8, 0.2) -- (0.0*8, -0.2) -- cycle;
  \draw (1.0*8, 0.2) -- (1.0*8, -0.2) -- cycle;
  \draw (2.0*8, 0.2) -- (2.0*8, -0.2) -- cycle;
  \node (0pointlabel) at (0.0,-0.5) {0};
  \node (1pointlabel) at (1.0*8,-0.5) {1};
  \node (2pointlabel) at (2.0*8,-0.5) {2};
  \node (xaxislabel) at (2.7*8,-0.5) {\textsc{SonAngle}};
  % dividing line
  \node (dividinglinestart) at (1.5 * 8,2) {};
  \node (dividinglineend)   at (1.5 * 8,-0.3) {};
  \path [line,dashed] (dividinglinestart) -- (dividinglineend) -- cycle;
  \node (dividingpointlabel) at (1.5*8, -0.5) {1.5};
\end{tikzpicture}
}

@@

The only clusters that give us evidence of this are /\textipa{Bm}/ clusters. These are not seen root-internally due to a high-ranking OCP effect \citep[cited in \cite{rose.2000}]{greenberg.1950}.
I suggest that there might be an undominated markedness constraint against this cluster -- either OCP(labial) or another constraint based on articulatory or perceptual difficulty -- that outranks all relevant faithfulness constraints.

@@

\newpage
\ex. {\sc *CCC} $\gg$ {\sc Dep}

\vspace{-2em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/CCC/} & *CCC & {\sc Dep} \\[0.5ex]
\hline \hline a. & & CCC & $\ast$! &  \\
\hline b. & \ding{43} & CC\textipa{1}C & & $\ast$ \\
\hline c. & \ding{43} & C\textipa{1}CC & & $\ast$ \\
\hline d. & & C\textipa{1}C\textipa{1}C & & $\ast\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\noindent {\bf [C\textipa{1}CC] idiolect.} In the third idiolect \ref{C1CCidiolect} (c.f. Section 5.2 of \cite{rose.2000}), most clusters epenthesise as [C\textipa{1}CC].

@@

Several possibilities are possible: retain the sonority of the underlying C\textsubscript{1} for that of the output value; use the sonority of the output C\textsubscript{1} as the underlying value; or take the angle formed by the meeting of the two contours, wherever that occurs.
